A user-extensible and adaptable parser architecture

### 2008-06-01 papers/sgai-2008/paper.tex
Addressed most of the remainder of Carl's comments. Still need to add examples, if possible.

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code (significantly lowering the barrier to entry for
new or casual users who need to parse new log lines), and greatly
simplifies the framework, actions and rules. Decoupling the actions from
the framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to perform any tasks specific to
the input being parsed; actions benefit from having services provided by
the framework, freeing them to concentrate on the task of accurately and
correctly processing the information provided by rules.

@@

The architecture's greatest strength is the ease with which is can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required when writing rules.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions: e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one or version two depending on the ruleset selected.
The architecture makes it possible to apply commonly used programming
techniques such as object orientation, inheritance, composition, delegation
and roles when designing and implementing a parser, simplifying the process
of working within a team with shared responsibility or developing and
testing additional functionality.

Developing a new parser based on this architecture follows a similar
process to developing a new parser using any architecture: study the
inputs, figure out what needs to happen for each category of input
(actions), and write a specification identifying each type of input
(rules); the existing framework should be suitable for use with the
majority of parsers.  If the inputs to be parsed are suitable a large
amount of work can be done with very little effort by using SLCT to
identify similar lines in the input and generate regexes to match those
lines; these regexes will form the basis of the initial ruleset.  When the
commonly occurring inputs are parsed by rules and actions, the process of
identifying and rectifying parsing deficiencies can begin, with each
iteration of the development process improving the parser, similar to
Stepwise Refinement~\cite{stepwise-refinement}.  This architecture is
ideally suited to parsing inputs where the input is not fully understood or
does not follow a fixed grammar: the architecture warns about unparsed
inputs and other errors, but continues on parsing as best it can, allowing
the developer of a new parser to decide which deficiencies are most
important and require attention first, rather than being forced to fix the
first error which arises.

>>>>

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code, significantly lowering the barrier to entry for
casual users who need to parse new log lines, e.g.\ part-time systems
administrators attempting to reduce spam.  Decoupling the actions from the
framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to perform any tasks specific to
the input being parsed; actions benefit from having services provided by
the framework, freeing them to concentrate on the task of accurately and
correctly processing the information provided by rules.

@@

The architecture's greatest strength is the ease with which is can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required when writing rules.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions: e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one or version two depending on the ruleset selected.
The architecture makes it possible to apply commonly used programming
techniques such as object orientation, inheritance, composition, delegation
and roles when designing and implementing a parser, simplifying the process
of working within a team with shared responsibility or developing and
testing additional functionality.  This architecture is ideally suited to
parsing inputs where the input is not fully understood or does not follow a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error which
arises.

The dataset gathered by the parser provides the foundation for the future
of this project: applying machine-learning algorithms to the data to
analyse and optimise the set of rules in use, followed by identifying
patterns in the data which could be used to write new filters to recognise
and reject spam rather than accepting it.  The current focus is on
clustering and decision trees to optimise the order in which rules are
applied; future efforts will involve using gathered data to train and test
new filters.  This task is very similar to optimising a black box
application based on its inputs and outputs, and this approach could be
applied to optimising the behaviour of any system given sufficient logging
to analyse.  An alternative approach to black box optimisation which uses
application profiling in conjunction with the application's error messages
to improve the error messages shown to users is described
in~\cite{black-box-error-reporting}; profiling data may be useful in
supplementing systems which fail to provide adequate log messages.
